A comparison on R tools for nonlinear least squares modeling

### 2023-03-06 RefactoringNLS/ComparingNLStools.Rmd
more edits

- handles formulas that include R functions, even ones which call calculations in
  other programming languages
- allows data to be weighted or subset
- can estimate bound constrained parameters
- provides a mechanism for handling partially linear models
- permits parameters to be indexed over a set of related data
- produces measures of variability (i.e., standard error estimates) for the 
estimated parameters
- has related profiling capabilities for exploring the likelihood surface as 
parameters are changed
- links to a number of pre-coded (`selfStart`) models.

@@

@@

Tools that produce a class `nls` output object create a rich set of functions
and structures that are then used in a variety of modeling tasks. By contrast,
`nlsr` computes quantities as they are requested or needed, with additional
features in separate functions. These two approaches
lead to different consequences for performance and how features are
provided. `nlsr` has antecendants in the methods of @cnm79, where storage for
data and programs was at a ridiculous premium in the small computers of the era.
Thus the code in `nlsr` is likely of value for workers to copy and modify it
for customized tools.

@@

Gauss-Newton/Marquardt methods all need a Jacobian matrix at each iteration. 
By default, `nlsr::nlxb()` will try to evaluate this using analytic 
expressions using symbolic and automatic differentiation tools. 
When using formula specification of the model, `nls()`, `minpack.lm::nlsLM()`
and `gslnls::gsl_nls()` use a finite difference approximation to compute the
Jacobian, though `gsl_nls()` does have an option to attempt symbolic expressions.

When The Jacobian code as 
the "gradient" attribute of the output of the Jacobian function. This 
lets us embed the code 
for the Jacobian as this attribute of the **residual** function so that the call 
to `nlfb()` can be made with the same name used for 
both residual and Jacobian function arguments. This programming trick saves a 
lot of trouble for the package developer, but it can be a nuisance for users 
trying to understand the code.

Generally `nls()`, `nlsLM()` and `gsl_nls()` use a fairly 
simple forward difference approximation of derivatives for the Jacobian
from the base-R `stats` package called `numericDeriv()`. A central 
approximation can be specified in control parameters. 
Package `nlsr` provides, via appropriate calling syntax, four approximation 
options, with a further control `ndstep` for the size of the
step used in the approximation.

@@

Analytic Jacobian code can be provided to `nls()`, `nlsLM()` and `gsl_nls()`,
and most `selfStart` models that automatically provide starting parameters
also include such code. There is documentation in R
of `selfStart` models, but the construction of them is non-trivial. A number 
of such models are included with base R in `./src/library/stats/R/zzModels.R` file 
with package `nlraa` (@MiguezNLRAA2021) providing a richer set. 
These provide the Jacobian in the "gradient" attribute of the "one-sided" formula
that defines
each model, and these Jacobians may be the analytic forms.

@@

In considering `selfStart` models, we noted that the base-R function `SSlogis`
is intended to solve our example problem **Logistic3T**, but found initial 
parameters for `nls()` by using a completely separate solver, then rerunning the
solution with the default algorithm from the derived -- and final -- parameters.
To provide simpler estimates of starting parameters, the function `SSlogisJN`
is now part of package `nlsr`.

@@

@@

@@

```{r ex10m, echo=TRUE}
# Hobbsmaskx.R -- masks with formula specification of problem
require(nlsr)
require(minpack.lm)
stu <- c(b1=200, b2=50, b3=0.3) # a default starting vector (named!)
sts <- c(c1=2, c2=5, c3=3) # a default scaled starting vector (named!)
traceval<-FALSE
# fix first parameter
anxbmsk1 <- try(nlxb(frmu, start=stu, data=weeddf, lower=c(200,0,0), 
			upper=c(200, 60, .3), trace=traceval))
print(anxbmsk1)
anlM1 <- try(nlsLM(frmu, start=stu, data=weeddf, lower=c(200,0,0), 
			upper=c(200, 60, .3), trace=traceval))
summary(anlM1)
# nls gives warnings
anlsmsk1 <- try(nls(frmu, start=stu, data=weeddf, lower=c(200,0,0), 
		upper=c(200, 60, .3),  algorithm="port", trace=traceval))
summary(anlsmsk1)
# Hobbs scaled problem with bounds, formula specification
anlxmsks1 <- nlxb(frms, start=sts, data=weeddf, lower=c(2,0,0),
                  upper=c(2,6,3))
print(anlxmsks1)
anlshmsk1 <- nls(frms, start=sts, trace=traceval, data=weeddf, lower=c(2,0,0),
             upper=c(2,6,3), algorithm='port')
summary(anlshmsk1)
cat("More precisely...crossprod(resid(anlshmsk1))=",crossprod(resid(anlshmsk1)),"\n")
# nlsLM does not always work with bounds
anlsLMmsks1 <- nlsLM(frms, start=sts, data=weeddf, lower=c(2,0,0),
                 upper=c(2,6,3))
summary(anlsLMmsks1)

@@

`nlsr` has an output format that indicates the constraint status of the parameter
estimates. For `nlsr` we have **chosen** to suppress calculation of 
approximate standard 
errors in the parameters when constraints are active, though we believe this 
policy worthy of discussion and further investigation.

@@

All four major mentioned must solve some variant of the Gauss-Newton 
equations. `nls()` uses a modification of an approach suggested by @Hartley1961,
while `nlsr`, `gslnls` and `minpack.lm` use flavours of @Marquardt1963. `gslnls`
offers an accelerated Marquardt method and three alternative methods. We have 
not explored all five methods with various control settings.
Though it is unlikely to be useful in general, 
control settings for `nlxb()` or `nlfb()` allow for those routines to perform 
a variety of Hartley and Marquardt algorithms. This
has been useful for learning about the algorithms, but tangential to simply finding
parameters of nonlinear models.

@@

While fixed weights may be useful, there are a number of problems for which
we want weights that are determined at least partially from the model 
parameters, for
example, a measure of the standard deviation of observations.

Such dynamic weighting situations are discussed in the vignette 
"Introduction to nlsr" of package `nlsr` in section 
*Weights that are functions of the model parameters*. Care is advised
in applying such ideas.

## Feature: Jacobian computation

`nls()`, with the `numericDeriv()` function, computes the Jacobian as the "gradient"
attribute of the residual vector. This is implemented as a mix of R and C code, but
we have created a rather more compact version entirely in R. 
See *https://github.com/nashjc/RNonlinearLS/tree/main/DerivsNLS*.
<!-- Code project. See the companion document **Jacobian Calculations for nls()** -->
<!-- (?? need citation -- currently **DerivsNLS.pdf**). -->

As far as we can understand the logic in `nls()`, the Jacobian computation during
parameter estimation is carried out entirely within the called C-language program,
and the R code function `numericDeriv()`, part of `./src/library/stats/R/nls.R`
in the R distribution source code. This is used to provide Jacobian information in 
the `nlsModel()` and `nlsModel.plinear()` functions, which are **not** exported for 
general use.

The Jacobian used by `nlsr::nlxb()` is, by default, computed from analytic expressions, 
while that
of `nls()` mostly uses a numerical approximation. Some `selfStart` models do provide
"gradient" information. `minpack.lm::nlsLM()` invokes `numericDeriv()` in its local 
version of `nlsModel()`, but it appears to use an internal approximate Jacobian code from 
the original Fortran `minpack` code, namely, `lmdif.f`. Such differences in approach
can lead to different behaviour, though in our experience differences are usually minor.

- A pasture regrowth problem (@Huet2004, page 1, based on @Ratkowsky1983) has a poorly
  conditioned Jacobian and `nls()` fails with "singular gradient". 
  Worse, numerical approximation to the Jacobian give the error
  "singular gradient matrix at initial parameter estimates" for `minpack.lm::nlsLM`
  so that the Marquardt stabilization is unable to take effect, while the analytic
  derivatives of `nlsr::nlxb` give a solution.

- Karl Schilling (private communication) provided an example where a model specified
  with the formula `y ~ a * (x ˆ b)` causes `nlsr::nlxb` to fail because the partial
  derivative w.r.t. `b` is `a * (x^b * log(x))`. If there is data for which `x = 0`,
  this is undefined. In such cases, we observed that `nls()` and `minpack.lm::nlsLM`
  found a solution, though it can be debated whether such lucky accidents can be
  taken as an advantage.

## Feature: missing start vector and self-starting models

Nonlinear estimation algorithms are almost all iterative and need a set of starting
parameters. `nls()` offers a special class of modeling functions called **`selfStart`** 
models. There are a number of these in base R (`./src/library/stats/R/zzModels.R`)
and others in R packages such as CRAN package `nlraa` (@MiguezNLRAA2021), as well as 
the now-archived package `NRAIA`. Unfortunately, some **`selfStart`** codes
entangle the calculation of starting values for parameters
with the particulars of the `nls()` code. Though there is a `getInitial()` function,
this is not easy to use to simply compute the initial parameter estimates outside of
`nls()`, in part 
because it may call that function. Such circular references are, in our view, dangerous.
Moreover, we believe that it would be helpful to have `selfStart` models that allow
users to explicitly provide values other than those suggested for the starting parameters.

Other concerns are revealed by the example below. Here, the `SSlogis` `selfStart` function 
is used to generate a set
of initial parameters for a 3-parameter logistic curve. The form used by `SSlogis`
is $y \,\sim\, Asym\,/\, (1 \,+\, exp((xmid\,-\,tt)\,/\,scal))$, but we show how the starting parameters
for this model can be transformed to those of 
another form of the model, namely, $y \,\sim\, b1\,/\,(1 \,+\, b2\,*\,exp(-b3\,*\,t))$.

The code for `SSlogis()` is in `./src/library/stats/R/zzModels.R`.

This R function includes analytic expressions for the Jacobian ("gradient").
These could be useful to R users, especially if documented. Moreover, we
wonder why the programmers have chosen to save so many quantities in "hidden"
variables, i.e., with names preceded by ".". These are then not displayed by the `ls()`
command, making them difficult to access.

### Strategic issues in `selfStart` models

By providing starting values that are likely to be reasonable and by including 
Jacobian code in the `selfStart` model code, some of the deficiencies of the 
Gauss-Newton algorithm are mitigated. However, creating such functions is a lot 
of work, and their
documentation (file `./src/library/stats/man/selfStart.Rd`) is quite complicated.
We believe that the focus should be placed on getting good initial parameters,
that is `getInitial()` function, though avoiding the current calls back to `nls()`.
Interactive tools, such as "visual fitting" (@nash1996nonlinear) might be
worth considering.

We also note that the introduction of `scaleOffset` in R 4.1.1 to deal with the 
convergence test for small residual problems now requires that the `getInitial()`
function have dot-arguments (`...`) in its argument list. This illustrates the
entanglement of many features in `nls()`.

## Issue: returned output of nls() and its documentation

The `nls` object contains some elements that are awkward to produce by other algorithms,
but some information that would be useful is not presented in a clear manner. Moreover, the
complexity of the object is a challenge to users.

In the following, we use `result` as the returned object from `nls()`.

The `data` return element is an R symbol. To actually access the data from this 
element, we need to use the syntax:

and if there is an element called `xdata`, it can be accessed as `result$model$xdata`.

Let us compare the `nls()` result with that from 
`nlsr::nlxb()`, which ostensibly solves the same problem:

```
 $ lower       : num [1:3] -Inf -Inf -Inf
 $ upper       : num [1:3] Inf Inf Inf
 $ maskidx     : int(0) 
 $ weights     : NULL
 $ formula     :Class 'formula'  language y ~ Asym/(1 + exp((xmid - tt)/scal))
  .. ..- attr(*, ".Environment")=<environment: R_GlobalEnv> 
 - attr(*, "class")= chr "nlsr"
```
The result object from `nlsr::nlxb()` is still much smaller than the one `nls()` returns. 
Moreover, `nlxb`
explicitly returns the sum of squares as well as the residual vector and Jacobian. 
The counts of evaluations are also returned.
(Note that the singular values of the Jacobian are
actually computed in the `print` and `summary` methods for the result.) 
As we made our comparisons, we noted several potential updates to the `nlsr` documentation
as well as that for `nls()`.

@@

As already noted, the function `resid()` (an alias for `residuals()`) is WEIGHTED, as are those in the
structured object `m` returned by `nls()` or `minpack.lm::nlsLM`,
e.g., `ans$m$resid()`. The function `nlsModel()`, which we have had to extract from the 
base R code and explicitly `source()` because it is not exported to the working namespace,
allows us to compute residuals for particular coefficient sets.

@@

@@

## Feature: partially linear models and their specification

The variable projection method (@Golub1973, @OlearyRust13) is generally much more 
effective than general approaches in finding good solutions to nonlinear least squares 
problems when some of the parameters appear linearly. However, setting up the 
calculations, that is, identifying which parameters are linear, is not a trivial
task.

Moreover, specifying a model to a solver should, ideally, use the same syntax across
solver tools. Unfortunately, R allows multiple approaches within different
modeling tools, and within `nls()` itself. This is highlighted by the case of
partially linear models.

The nonlinear modeling specifications are, of course, a development of linear ones.
Unfortunately, the explicit model `y ~ a * x + b` does not work with the linear modeling 
function `lm()`, which requires this model to be specified as `y ~ x`.

Within `nls()`, consider the following FOUR different specifications for the same problem, plus 
an intuitive choice, labelled `fm2a`, that does NOT work. In this failed
attempt, putting the `Asym` parameter in the model causes the `plinear` algorithm
to try to add another term to the model. We believe this is unfortunate, and would
like to see a consistent syntax. At the time of writing, we do
not foresee a resolution for this issue. In the example, we have NOT evaluated
the commands to save space.

@@

## Feature: indexed parameters

@@

## Feature: bounds on parameters

There are many situations where the context of a problem constrains the values of parameters.
For example, one of us was asked many years ago to estimate a model where one parameter estimate
was negative. The client complained "But that is supposed to be the number of grain elevators
in Saskatchewan". The number should have been a positive integer, and likely less
than a few thousand.

`nls()` can impose bounds on parameters, but only if the `port` algorithm is used. Unfortunately,
the manual states

> The `algorithm = "port"` code appears unfinished, and does not even check that the starting 
  value is within the bounds. Use with caution, especially where bounds are supplied.

This is clearly unsatisfactory for general use, but does find bounded solutions. 
Inclusion of bounds on parameters in nonlinear least squares computations is relatively 
straightforward, and it is part of the default method
for package `nlsr`. Package `minpack.lm` also includes provision for bounds, but the following
script, which we do not run here for reasons of space, shows that it does not find a
good solution for the scaled Hobbs test problem.

>>>>

- handles formulas that include R functions, even ones which call calculations in
  other programming languages
- allows data to be weighted or subset
- can estimate bound constrained parameters
- provides a mechanism for handling partially linear models
- permits parameters to be indexed over a set of related data
- produces measures of variability (i.e., standard error estimates) for the 
estimated parameters
- has related profiling capabilities for exploring the likelihood surface as 
parameters are changed
- links to a number of pre-coded (`selfStart`) models that do not require
  initial parameter values.

@@

### Returned results of nls() and other tools

The complexity of this object is a challenge to users. Let us use 
`result` as the returned object from `nls()`, `minpack.lm::nlsLM()` or
`gslnls::gsl_nls()`. For example, the `data` return element is an R symbol. To
actually access the data from this element, we need to use the syntax:

and if there is an element called `xdata` that can be accessed as
`result$model$xdata`.

By contrast `nlsr::nlxb()` returns

```
 $ lower       : num [1:3] -Inf -Inf -Inf
 $ upper       : num [1:3] Inf Inf Inf
 $ maskidx     : int(0) 
 $ weights     : NULL
 $ formula     :Class 'formula'  language y ~ Asym/(1 + exp((xmid - tt)/scal))
  .. ..- attr(*, ".Environment")=<environment: R_GlobalEnv> 
 - attr(*, "class")= chr "nlsr"
```
The result object from `nlsr::nlxb()` is still much smaller than the one `nls()` 
returns. Moreover, `nlxb`
explicitly returns the sum of squares as well as the residual vector and Jacobian. 
The counts of evaluations are also returned.

@@

Tools that produce a class `nls` output object create a rich set of functions
and structures that are then used in a variety of modeling tasks. By contrast,
`nlsr` computes quantities as they are requested or needed, with additional
features in separate functions. For example, the singular values of the Jacobian 
are actually computed in the `print` and `summary` methods for the result. 
These two approaches
lead to different consequences for performance and how features are
provided. `nlsr` has antecendants in the methods of @cnm79, where storage for
data and programs was at a ridiculous premium in the small computers of the era.
Thus the code in `nlsr` is likely of value for workers to copy and modify it
for customized tools.

@@

Gauss-Newton/Marquardt methods all need a Jacobian matrix at each iteration. 
By default, `nlsr::nlxb()` will try to evaluate this using analytic 
expressions using symbolic and automatic differentiation tools. 
When using a formula specification of the model, `nls()`, `minpack.lm::nlsLM()`
and `gslnls::gsl_nls()` use a finite difference approximation to compute the
Jacobian, though `gsl_nls()` does have an option to attempt symbolic expressions.
Package `nlsr` provides, via appropriate calling syntax, four numeric approximation 
options for the Jacobian, with a further control `ndstep` for the size of the
step used in the approximation.

Using the "gradient" attribute of the output of the Jacobian function to hold
this matrix lets us embed this attribute of the **residual** function as well,
so that the call to `nlfb()` can be made with the same name used for 
both residual and Jacobian function arguments. This programming trick saves a 
lot of trouble for the package developer, but it can be a nuisance for users 
trying to understand the code.

As far as we can understand the logic in `nls()`, the Jacobian computation during
parameter estimation is carried out entirely within the called C-language program,
and the R code function `numericDeriv()`, part of `./src/library/stats/R/nls.R`
in the R distribution source code. This is used to provide Jacobian information in 
the `nlsModel()` and `nlsModel.plinear()` functions, which are **not** exported for 
general use. `gsl_nls()` also appears to use `numericDeriv()`.

`numericDeriv()` uses a 
simple forward difference approximation of derivatives, though a central 
difference approximation can be specified in control parameters. 
We are unclear why `numericDeriv()` in base R calls `C_numeric_deriv`,
as we were easily able to create a more compact version entirely in R. 
See *https://github.com/nashjc/RNonlinearLS/tree/main/DerivsNLS*.

`minpack.lm::nlsLM()` invokes `numericDeriv()` in its local 
version of `nlsModel()`, but it appears to use an internal approximate Jacobian 
code from 
the original Fortran `minpack` code, namely, `lmdif.f`. Such differences in approach
can lead to different behaviour, though in our experience the differences are usually
very minor.

Minor numerical differences can give annoying results with ill-conditioned
problems.

- A pasture regrowth problem (@Huet2004, page 1, based on @Ratkowsky1983) has a
  poorly conditioned Jacobian and `nls()` fails with "singular gradient". 
  Worse, numerical approximation to the Jacobian gives the error
  "singular gradient matrix at initial parameter estimates" for `minpack.lm::nlsLM`
  so that the Marquardt stabilization is unable to take effect, while the analytic
  derivatives of `nlsr::nlxb` give a solution.

- Karl Schilling (private communication) provided an example where a model specified
  with the formula `y ~ a * (x ˆ b)` causes `nlsr::nlxb` to fail because the partial
  derivative w.r.t. `b` is `a * (x^b * log(x))`. If there is data for which `x = 0`,
  the derivative is undefined, while the model can be computed. In such cases, 
  we observed that `nls()` and 
  `minpack.lm::nlsLM` found a solution, though this seems to be a lucky accident.

@@

Analytic Jacobian code can be provided to `nls()`, `nlsLM()` and `gsl_nls()`,
and most `selfStart` models that automatically provide starting parameters
also include such code. There is documentation in R
of `selfStart` models, but the construction of them is non-trivial. A number 
of such models are included with base R in `./src/library/stats/R/zzModels.R` file 
with package `nlraa` (@MiguezNLRAA2021) providing a richer set. 
There are also some in the now-archived package `NRAIA`.
These provide the Jacobian in the "gradient" attribute of the "one-sided" formula
that defines
each model, and these Jacobians may be the analytic forms.

@@

In considering `selfStart` models, we noted that the base-R function `SSlogis`
is intended to solve our example problem **Logistic3T**, but that it finds initial 
parameters for `nls()` by using a completely separate solver, then reruns the
solution with the default algorithm from the derived -- and final -- parameters.
To provide simpler estimates of starting parameters, the function `SSlogisJN`
is now part of package `nlsr`.

The `SSLogis` case is one where the `getInitial()` function (used to find
starting values) actually calls `nls()` with a non-default algorithm. 
Such circular references are, in our view, prone to creating errors
in code maintenance.
Users may also want to use these models with explicit starting values 
other than those suggested by `getInitial()`.
We are also surprised that the 
analytic expressions for the Jacobian ("gradient") in the `SSLogis` R function 
saves many quantities in "hidden"
variables, i.e., with names preceded by ".". 
These are then not displayed by the `ls()`
command, making them difficult to access by users who may wish to create
their own selfStart model by a copy and edit.
Interactive tools, such as "visual fitting" (@nash1996nonlinear) might be
worth considering as a way to find starting parameters, but we know of no
R capability of this type.

As a side note, the introduction of `scaleOffset` in R 4.1.1 to deal with the 
convergence test for small residual problems now requires that the `getInitial()`
function have dot-arguments (`...`) in its argument list. This illustrates the
entanglement of many features in `nls()` that complicate its maintenance and
improvement.

@@

?? Arkajyoti -- we can probably trim these examples

@@

- bounds are part of the default method for package `nlsr`.

@@

```{r ex10m, echo=TRUE}
# Hobbsmaskx.R -- masks with formula specification of problem
require(nlsr)
require(minpack.lm)
stu <- c(b1=200, b2=50, b3=0.3) # a default starting vector (named!)
sts <- c(c1=2, c2=5, c3=3) # a default scaled starting vector (named!)
traceval<-FALSE
# fix first parameter
anxbmsk1 <- try(nlxb(frmu, start=stu, data=weeddf, lower=c(200,0,0), 
			upper=c(200, 60, 3), trace=traceval))
print(anxbmsk1)
anlM1 <- try(nlsLM(frmu, start=stu, data=weeddf, lower=c(200,0,0), 
			upper=c(200, 60, 3), trace=traceval))
summary(anlM1)
# nls gives warnings
anlsmsk1 <- try(nls(frmu, start=stu, data=weeddf, lower=c(200,0,0), 
		upper=c(200, 60, 3),  algorithm="port", trace=traceval))
summary(anlsmsk1)
# Hobbs scaled problem with bounds, formula specification
anlxmsks1 <- nlxb(frms, start=sts, data=weeddf, lower=c(2,0,0),
                  upper=c(2,6,30))
print(anlxmsks1)
anlshmsk1 <- nls(frms, start=sts, trace=traceval, data=weeddf, lower=c(2,0,0),
             upper=c(2,6,3), algorithm='port')
summary(anlshmsk1)
cat("More precisely...crossprod(resid(anlshmsk1))=",crossprod(resid(anlshmsk1)),"\n")
# nlsLM does not always work with bounds
anlsLMmsks1 <- nlsLM(frms, start=sts, data=weeddf, lower=c(2,0,0),
                 upper=c(2,6,3))
summary(anlsLMmsks1)

@@

`nlsr` has an output format that indicates the constraint status of the parameter
estimates. For `nlsr` we have **chosen** to suppress calculation of 
approximate standard 
errors in the parameters when constraints are active because their meaning
under constraints is unclear, though we believe this 
policy worthy of discussion and further investigation.

@@

All four major tools illustrated must solve some variant of the Gauss-Newton 
equations. `nls()` uses a modification of an approach suggested by @Hartley1961,
while `nlsr`, `gslnls` and `minpack.lm` use flavours of @Marquardt1963. `gslnls`
offers an accelerated Marquardt method and three alternative methods. We have 
not explored all five methods with various control settings.
Though it is unlikely to be useful in general, 
control settings for `nlxb()` or `nlfb()` allow for those routines to perform 
a variety of Hartley and Marquardt algorithms. This
has been useful for learning about the algorithms, but tangential to simply finding
parameters of nonlinear models.

@@

While fixed weights may be useful, there are a number of problems for which
we want weights that are determined at least partially from the model 
parameters, for
example, a measure of the standard deviation of observations.
Such dynamic weighting situations are discussed in the vignette 
"Introduction to nlsr" of package `nlsr` in section 
*Weights that are functions of the model parameters*. Care is advised
in applying such ideas.

@@

The function `resid()` (an alias for `residuals()`) gives WEIGHTED residuals, 
as does, for example, `result$m$resid()`. The function `nlsModel()`, which we 
have had to extract from the base R code and explicitly `source()` because it 
is not exported to the working namespace,
allows us to compute residuals for particular coefficient sets.

@@

## Minor complaints

@@

## Estimating models that are partially linear

Many nonlinear models can be expressed so that some of the parameters 
appear linearly. In our logistic examples, the asymptote parameters are
an illustration. In such cases, specialized methods can give spectactularly
better performance. The variable projection method (@Golub1973, @OlearyRust13) is 
generally much more effective than general approaches in finding good solutions 
to nonlinear least squares 
problems when some of the parameters appear linearly. However, setting up the 
calculations, that is, identifying which parameters are linear, is not a trivial
task.

`nls()` has an option `algorithm="plinear"` that allows some partially linear
models to be solved. The other tools, as far as we are aware, do not offer any
such capability, with no external package providing it. 
The feature in `nls()` has, however, some awkwardness.
In particular, we would like to specify a model to a solver consistently
across solver methods. Unfortunately, within `nls()` itself we must use
different specifications with different `algorithm` options.

The nonlinear model specifications are, of course, a development of linear ones.
Unfortunately, the explicit model `y ~ a * x + b` does not work with the linear
modeling function `lm()`, which requires this model to be specified as `y ~ x`.
Within `nls()`, consider the following FOUR different specifications for the same 
problem, plus an intuitive choice, labelled `fm2a`, that does NOT work. 
In this failed attempt, putting the `Asym` parameter in the model causes the
`plinear` algorithm
to try to add another term to the model. We believe this is unfortunate, and would
like to see a consistent syntax. At the time of writing, we do
not foresee a resolution for this issue. In the example, we have NOT evaluated
the commands to save space.

@@

## Models with indexed parameters

@@

If such models are of enough importance, we believe it would be relatively
straightforward to prepare an external package to provide for their estimation,
including an appropriately formatted output object so the indexing of the 
parameters can be used in other calculations.
